Type Classes for Mathematics in Type Theory

### 2010-06-28 jar.tex
Add paragraph about implicit generalization and the functionality we would like to see.

% todo:
  % cite that other type class based development in the context of the alternative bundling schemes.
  % maybe have a chapter called "interfaces are primary, not implementations" (also involve polynomials)

@@

% todo: mention proof irrelevance here somewhere, and terms that depend on proofs.

@@

A remaining eyesore in the theorem declaration is the enumeration of \lstinline|e| and \lstinline|op|. To remove these, we use a new parameter declaration feature called \emph{implicit generalization}, introduced in Coq specifically to support type classes. Using implicit generalization, we can write:

\begin{lstlisting}
  Lemma L `{SemiGroup G}: Π x y z: G, x & y & z = z & y & x.
\end{lstlisting}
% TODO: describe implicit generalization in two words
% TODO: at some point down the line in our coq-inadequacies section, describe the implicit generalization improvements we need.
Thus, we have reached the mathematical ideal we aimed for.

>>>>

% todo:
  % cite that other type class based development in the context of the alternative bundling schemes.

@@

@@

A remaining eyesore in the theorem declaration is the enumeration of \lstinline|e| and \lstinline|op|. To remove these, we use a new parameter declaration feature called \emph{implicit generalization}, introduced in Coq specifically to support type classes. Using implicit generalization, we can write:
\begin{lstlisting}
  Lemma L `{SemiGroup G}: Π x y z: G, x & y & z = z & y & x.
\end{lstlisting}
The backtick tells Coq to insert implicit declarations of further parameters to \lstinline|SemiGroup G|, namely those declared as \lstinline|e| and \lstinline|op| above. It further lets us omit a name for the \lstinline|SemiGroup G| parameter itself as well. All of these will be given automatically generated names that we will never refer to.

Thus, we have reached the mathematical ideal we aimed for.

While we are on the topic of implicit generalization, we mention one inadequacy concerning their current implementation that we feel should be addressed for the facility to be a completely satisfying solution. While the syntax already supports variants (not shown above) that differ in how exactly different kinds of arguments are inferred and/or generalized, there is no support to have an argument be ``inferred if possible, generalized otherwise". We have found that the need for such a policy arises naturally when declaring a parameter of class type in a contexts where \emph{some} of its components are already available, while others are to be newly introduced. The current workaround in these cases involves providing names for components that are then never referred to, which is a bit awkward.
